My title

### 2023-03-15 outputs/paper/paper.qmd
final graph and data cleaning

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8)

### Input data ###
income_data_percentage<-read.csv(here::here("outputs/data/income_3years.csv"))
sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
race_data<-read.csv(here::here("outputs/data/race_data.csv"))
employee_status_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
industry_sector_data<-read.csv(here::here("outputs/data/industry_sector_data.csv"))
science_course_data<-read.csv(here::here("outputs/data/science_course_data.csv"))

@@

```{r, fig.height=8}
#| message: false
#| warning: false
#| echo: false
#| label: fig-1
#| fig-cap: "U.S. Income Distribution from 2014 to 2018"

# Make sure the order of the income level is the order we wanted
income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")

income_data_percentage$Income <- factor(income_data_percentage$Income, levels = income_label)

# Define color palette with 3 colors
fill_colors <- brewer.pal(n = 3, name = "RdYlBu")

# Generate graph
ggplot(income_data_percentage, aes(x = Income, y = Number, fill = as.factor(Year))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = fill_colors) +
  labs(x = "Income Level", y = "Total Number of People", title = "U.S. Income Distribution by Year") +
  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = Number + 10), position = position_stack(vjust = 0.5), size = 2) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  ylim(0, 2400)

@@

# Data {#sec-data}

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-2
#| fig-cap: "U.S. Income Distribution by Sex"

# calculate the percentage of each sex group for each year
sex_data <- sex_data %>% 
  group_by(year) %>% 
  mutate(percentage = 100 * people / sum(people))

# generate graph
ggplot(sex_data, aes(x = year, y = people, 
                                 fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 1), size = 3) +
  labs(x = "Year", y = "People", fill = "Sex",title = "sex")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-3
#| fig-cap: "U.S. Income Distribution by Race"

# calculate the percentage of each race group for each year
race_data <- race_data %>% 
  group_by(year) %>% 
  mutate(percentage = 100 * people / sum(people))

# generate graph
ggplot(race_data, aes(x = year, y = people, 
                                 fill = race)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 2), size = 3) +
  labs(x = "Year", y = "People", fill = "Race",title = "race")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-4
#| fig-cap: "U.S. Income Distribution by Employee Status"

# calculate the percentage of each employee status group for each year
employee_status_data <- employee_status_data %>% 
  group_by(year) %>% 
  mutate(percentage = 100 * people / sum(people))

# generate graph
ggplot(employee_status_data, aes(x = year, y = people, 
                                 fill = employee_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
  labs(x = "Year", y = "People", fill = "Employee Status",title = "employee status")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-5
#| fig-cap: "U.S. Income Distribution by Industry Sector"

# calculate the percentage of each industry sector group for each year
industry_sector_data <- industry_sector_data %>% 
  group_by(year) %>% 
  mutate(percentage = 100 * people / sum(people))

# generate graph
ggplot(industry_sector_data, aes(x = year, y = people, 
                                 fill = industry_sector)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
  labs(x = "Year", y = "People", fill = "Industry Sector",title = "industry sector")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6
#| fig-cap: "U.S. Income Distribution by College Level Science Courses"

# calculate the percentage of taking college level science courses for each year
science_course_data <- science_course_data %>% 
  group_by(year) %>% 
  mutate(percentage = 100 * people / sum(people))

# generate graph
ggplot(science_course_data, aes(x = year, y = people, 
                                 fill = science_course)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
  labs(x = "Year", y = "People", fill = "Whether taking science course",title = "Science course")

>>>>

@@

# Methodology

```{r setup, include=FALSE}}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8)

### Input data ###
happiness_data<-read.csv(here::here("outputs/data/happiness_data.csv"))
sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
race_data<-read.csv(here::here("outputs/data/race_data.csv"))
degree_data<-read.csv(here::here("outputs/data/degree_data.csv"))
employee_status_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))

@@

# Results
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-1Happiness
#| fig-cap: "Changes in happiness levels from 1972 to 2021"

# generate graph
happy_colors <- c("#FF7F00", "#1F78B4", "#33A02C")

ggplot(happiness_data, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = happy_colors) +
  scale_color_manual(values = happy_colors) +
  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
       title = "Changes in happiness levels from 1972 to 2021",
       color = "Happy level")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-2VeryHappyGender
#| fig-cap: "Trends in happiness level (very happy) by gender from 1972 to 2021"

# calculate the count of individuals for very happy level
sex_data_happy1 <- sex_data %>% 
  filter(Happy == "Very happy") %>% 
  group_by(Sex, Year, Happy) %>% 
  summarize(People = n())

# generate graph
sex_colors <- c("#E74C3C", "#2C3E50")

ggplot(sex_data_happy1, aes(x = Year, y = People, fill = Sex)) +
  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
  scale_fill_manual(values = sex_colors) +
  scale_color_manual(values = sex_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (very happy) by gender from 1972 to 2021")

```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-3PrettyHappyGender
#| fig-cap: "Trends in happiness level (pretty happy) by gender from 1972 to 2021"

# calculate the count of individuals for pretty happy level
sex_data_happy2 <- sex_data %>% 
  filter(Happy == "Pretty happy") %>% 
  group_by(Sex, Year, Happy) %>% 
  summarize(People = n())

# generate graph
sex_colors <- c("#E74C3C", "#2C3E50")

ggplot(sex_data_happy2, aes(x = Year, y = People, fill = Sex)) +
  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
  scale_fill_manual(values = sex_colors) +
  scale_color_manual(values = sex_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (pretty happy) by gender from 1972 to 2021")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-4NotTooHappyGender
#| fig-cap: "Trends in happiness level (not too happy) by gender from 1972 to 2021"

# calculate the count of individuals for not too happy level
sex_data_happy3 <- sex_data %>% 
  filter(Happy == "Not too happy") %>% 
  group_by(Sex, Year, Happy) %>% 
  summarize(People = n())

# generate graph
sex_colors <- c("#E74C3C", "#2C3E50")

ggplot(sex_data_happy3, aes(x = Year, y = People, fill = Sex)) +
  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
  scale_fill_manual(values = sex_colors) +
  scale_color_manual(values = sex_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (not too happy) by gender from 1972 to 2021")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-5VeryHappyRace
#| fig-cap: "Trends in happiness level (very happy) by race from 1972 to 2021"

# calculate the count of individuals for very happy level
race_data_happy1 <- race_data %>% 
  filter(Happy == "Very happy") %>% 
  group_by(Race, Year, Happy) %>% 
  summarize(People = n())

# generate graph
race_colors <- c("#3498DB", "#27AE60", "#F39C12")

ggplot(race_data_happy1, aes(x = Year, y = People, fill = Race)) +
  geom_line(aes(group = Race, color = Race), size = 1.2) +
  scale_fill_manual(values = race_colors) +
  scale_color_manual(values = race_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (very happy) by race from 1972 to 2021")

@@

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6PrettyHappyRace
#| fig-cap: "Trends in happiness level (pretty happy) by race from 1972 to 2021"

# calculate the count of individuals for pretty happy level
race_data_happy2 <- race_data %>% 
  filter(Happy == "Pretty happy") %>% 
  group_by(Race, Year, Happy) %>% 
  summarize(People = n())

# generate graph
race_colors <- c("#3498DB", "#27AE60", "#F39C12")

ggplot(race_data_happy2, aes(x = Year, y = People, fill = Race)) +
  geom_line(aes(group = Race, color = Race), size = 1.2) +
  scale_fill_manual(values = race_colors) +
  scale_color_manual(values = race_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (pretty happy) by race from 1972 to 2021")

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-7NotTooHappyRace
#| fig-cap: "Trends in happiness level (not too happy) by race from 1972 to 2021"

# calculate the count of individuals for not too happy level
race_data_happy3 <- race_data %>% 
  filter(Happy == "Not too happy") %>% 
  group_by(Race, Year, Happy) %>% 
  summarize(People = n())

# generate graph
race_colors <- c("#3498DB", "#27AE60", "#F39C12")

ggplot(race_data_happy3, aes(x = Year, y = People, fill = Race)) +
  geom_line(aes(group = Race, color = Race), size = 1.2) +
  scale_fill_manual(values = race_colors) +
  scale_color_manual(values = race_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (not too happy) by race from 1972 to 2021")

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-8LessHighSchool
#| fig-cap: "Changes in happiness levels of individuals with less than high school education from 1972 to 2021"

# calcualte the people who have less than high school degree based on different level of happy
degree_data_1 <- degree_data %>% 
  filter(Degree == "Less than high school") %>% 
  group_by(Year, Happy) %>% 
  summarize(People = n())

# generate graph
degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")

ggplot(degree_data_1, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = degree_colors) +
  scale_color_manual(values = degree_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith less than high school education from 1972 to 2021", color = "Happy level")

```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-9HighSchool
#| fig-cap: "Changes in happiness levels of individuals with high school education from 1972 to 2021"

# calcualte the people who have high school degree based on different level of happy
degree_data_2 <- degree_data %>% 
  filter(Degree == "High school") %>% 
  group_by(Year, Happy) %>% 
  summarize(People = n())

# generate graph
degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")

ggplot(degree_data_2, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = degree_colors) +
  scale_color_manual(values = degree_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith high school education from 1972 to 2021", color = "Happy level")

```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-10AssociateCollege
#| fig-cap: "Changes in happiness levels of individuals with associate/junior college education from 1972 to 2021"

# calcualte the people who have associate/junior college degree based on different level of happy
degree_data_3 <- degree_data %>% 
  filter(Degree == "Associate/junior college") %>% 
  group_by(Year, Happy) %>% 
  summarize(People = n())

# generate graph
degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")

ggplot(degree_data_3, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = degree_colors) +
  scale_color_manual(values = degree_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith associate/junior college education from 1972 to 2021", color = "Happy level")
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-11Bachelor
#| fig-cap: "Changes in happiness levels of individuals with bachelor's education from 1972 to 2021"

# calcualte the people who have Bachelor's degree based on different level of happy
degree_data_4 <- degree_data %>% 
  filter(Degree == "Bachelor's") %>% 
  group_by(Year, Happy) %>% 
  summarize(People = n())

# generate graph
degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")

ggplot(degree_data_4, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = degree_colors) +
  scale_color_manual(values = degree_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith bachelor's education from 1972 to 2021", color = "Happy level")

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-12Graduate
#| fig-cap: "Changes in happiness levels of individuals with graduate education from 1972 to 2021"

# calcualte the people who have graduate degree based on different level of happy
degree_data_5 <- degree_data %>% 
  filter(Degree == "Graduate") %>% 
  group_by(Year, Happy) %>% 
  summarize(People = n())

# generate graph
degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")

ggplot(degree_data_5, aes(x = Year, y = People, fill = Happy)) +
  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
  scale_fill_manual(values = degree_colors) +
  scale_color_manual(values = degree_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith graduate education from 1972 to 2021", color = "Happy level")

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-13EmployeeStatusVeryHappy
#| fig-cap: "Trends in happiness level (very happy) by employee status from 1972 to 2021"

# calcualte the number of employee status for very happy
employee_status_data_happy1 <- employee_status_data %>% 
  filter(Happy == "Very happy") %>% 
  group_by(Employee_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
employee_status_colors <- c("#E74C3C", "#2C3E50")

ggplot(employee_status_data_happy1, aes(x = Year, y = People, fill = Employee_status)) +
  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
  scale_fill_manual(values = employee_status_colors) +
  scale_color_manual(values = employee_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (very happy) by employee status from 1972 to 2021", color = "Employee status")

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-14EmployeeStatusPrettyHappy
#| fig-cap: "Trends in happiness level (pretty happy) by employee status from 1972 to 2021"

# calcualte the number of employee status for pretty happy
employee_status_data_happy2 <- employee_status_data %>% 
  filter(Happy == "Pretty happy") %>% 
  group_by(Employee_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
employee_status_colors <- c("#E74C3C", "#2C3E50")

ggplot(employee_status_data_happy2, aes(x = Year, y = People, fill = Employee_status)) +
  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
  scale_fill_manual(values = employee_status_colors) +
  scale_color_manual(values = employee_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (pretty happy) by employee status from 1972 to 2021", color = "Employee status")
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-15EmployeeStatusNotTooHappy
#| fig-cap: "Trends in happiness level (not too happy) by employee status from 1972 to 2021"

# calcualte the number of employee status for not to happy
employee_status_data_happy3 <- employee_status_data %>% 
  filter(Happy == "Not too happy") %>% 
  group_by(Employee_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
employee_status_colors <- c("#E74C3C", "#2C3E50")

ggplot(employee_status_data_happy3, aes(x = Year, y = People, fill = Employee_status)) +
  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
  scale_fill_manual(values = employee_status_colors) +
  scale_color_manual(values = employee_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (not too happy) by employee status from 1972 to 2021", color = "Employee status")
```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-16MaritalStatusVeryHappy
#| fig-cap: "Trends in Happiness Level (very happy) by Marital Status from 1972 to 2021"

# calcualte the number of marital status for very happy
marital_status_data_happy1 <- marital_status_data %>% 
  filter(Happy == "Very happy") %>% 
  group_by(Marital_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")

ggplot(marital_status_data_happy1, aes(x = Year, y = People, fill = Marital_status)) +
  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
  scale_fill_manual(values = marital_status_colors) +
  scale_color_manual(values = marital_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (very happy) by marital status from 1972 to 2021", color = "Marital status")
```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-17MaritalStatusPrettyHappy
#| fig-cap: "Trends in happiness level (pretty happy) by marital status from 1972 to 2021"

# calcualte the number of marital status for pretty happy
marital_status_data_happy2 <- marital_status_data %>% 
  filter(Happy == "Pretty happy") %>% 
  group_by(Marital_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")

ggplot(marital_status_data_happy2, aes(x = Year, y = People, fill = Marital_status)) +
  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
  scale_fill_manual(values = marital_status_colors) +
  scale_color_manual(values = marital_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (pretty happy) by marital status from 1972 to 2021", color = "Marital status")
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-18MaritalStatusNotTooHappy
#| fig-cap: "Trends in happiness level (not too happy) by marital status from 1972 to 2021"

# calcualte the number of marital status for not too happy
marital_status_data_happy3 <- marital_status_data %>% 
  filter(Happy == "Not too happy") %>% 
  group_by(Marital_status, Year, Happy) %>% 
  summarize(People = n())

# generate graph
marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")

ggplot(marital_status_data_happy3, aes(x = Year, y = People, fill = Marital_status)) +
  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
  scale_fill_manual(values = marital_status_colors) +
  scale_color_manual(values = marital_status_colors) +
  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (not too happy) by marital status from 1972 to 2021", color = "Marital status")
```

```{r}

```{r}

```{r}
